Week 14 Phylogenetic community analysis
========================================================

```{r}
require(picante)
tree<- read.tree("hum294.tre")
plot(tree)
```

```{r}
cophen<-cophenetic(tree2)
hist(cophen)
```

```{r}
sitesp<-read.csv("SiteXspp.csv")
```

We need to do a bunch of crap to reformat the species names and remove accession numbers and make them consistent between the phylogeny and the site by species matrix

something something

```{r}
vec<-strsplit(tree$tip.label,"\\.")
for(i in 1:nrows(tree){
  paste(vec[1:2])
  #etc etc
}
```

But in the meantime, reading in the tree that John fixed faster than I did

```{r}
tree2<-read.tree("newtree.tre")
plot(tree2)
```

```{r}
cophen<-cophenetic(tree2)
hist(cophen)
```

```{r}
sitesp<-read.csv("SiteXspp.csv", row.names=1)
```

check that the site x sp matrix and the tree names match

```{r}
colnames(sitesp[,2:ncol(sitesp)]) %in% tree2$tip.label
tree2$tip.label %in% colnames(sitesp[,2:ncol(sitesp)])

#counting things in the first that aren't in the second
sum(!colnames(sitesp[,2:ncol(sitesp)]) %in% tree2$tip.label)
sum(!tree2$tip.label %in% colnames(sitesp[,2:ncol(sitesp)]))
```

There are 40 species in the species matrix that aren't in our tree.
Get the matrix and the tree to match...

This bit of code below subsamples the matrix (don't run this directly, we just need this to run the MPD code below)
```{r}
sitesp[,colnames(sitesp) %in% tree2$tip.label]
```

Calculating mean phylogenetic distance (a test statistic that describes each site)
```{r}
mpd<-mpd(sitesp[,colnames(sitesp) %in% tree2$tip.label], cophen)
```

Make a data frame that lists the site names and the mpd

```{r}
MPDtable<-as.data.frame(mpd)
row.names(MPDtable)<-row.names(sitesp)
```


Randomize the site by species matrix and resample MPD 1000 times  (make a function that appends test statistics to the table you already have - site, obserbed mpd, mpdrand1, mpdrand2, etc...)  r00 is a totally random new community

```{r}
require(vegan)

for (i in 1:100){
randcom = commsimulator(sitesp, "r00")
randmpd = mpd(randcom[,colnames(randcom) %in% tree2$tip.label], cophen)
MPDtable = cbind(MPDtable, randmpd)
}
```


Repeat the process for r0 (retains row sums) and c0 (retains row and column sums)
