Week 2 Exercises
============
Emily Rollinson and Sarah Supp. 


1) Decide on the best distribution for the clutch size data

```{r}
clutch<-read.csv("ClutchSize.csv")
head(clutch)
mean<-mean(clutch$Clutch_size)
sd<-sd(clutch$Clutch_size)
mean
sd
hist(log(clutch$Clutch_size))
```
The distribution appears log-normal.

2) Using the statistic <i>mean clutch size</i>, compare the results of bootstrapping 3 ways:  


a) sample with replacement from species regardless of higher taxonomic information; 

```{r}

mean(clutch$Clutch_size)   
n = nrow(clutch)

est = NULL
for (i in 1:1000){
  boot = sample(clutch$Clutch_size, n, replace=TRUE)
  mean = mean(boot)
  est = append(est,mean)
}     

      mean(est)
```

b) Sample with replacement from the list of families, then sample the species from that list with replacement and bootstrap the species samples.

```{r}
      f=length(unique(clutch$Family))
est2 = NULL
for (i in 1:1000){
  fam = sample(clutch$Family, f, replace = TRUE)
  subdat = clutch[which(clutch$Family %in% fam),]
  boot = sample(subdat$Clutch_size, n, replace=TRUE)
  mean = mean(boot)
  est2 = append(est,mean)
}     

      mean(est2)
      
```



c) Sample families with replacement.  Within that set of samples, sample genera with replacement.  Within this second set of samples, sample species with replacement, and create a bootstrap estimate of clutch size mean from this sampled list of species.  This may not be exactly right.

```{r}
      f=length(unique(clutch$Family))
est3 = NULL
for (i in 1:1000){
  fam = sample(clutch$Family, f, replace = TRUE)
  subdat = clutch[which(clutch$Family %in% fam),]
   g=length(unique(subdat$Genus_name))
  gen = sample(subdat$Genus_name, g, replace = TRUE)
  subdat2 = subdat[which(subdat$Genus_name %in% gen),]
  boot = sample(subdat2$Clutch_size, n, replace=TRUE)
  mean = mean(boot)
  est3 = append(est3,mean)
}     

mean(est3)
      
```