Sam's Adventures in Heirarchical Bootstrapping
========================================================

```{r, message=FALSE, warning=FALSE}
library(MASS) # for fitdistr
library(VGAM) # for rayleight distribution
clutch <- read.csv("ClutchSize.csv")
````

Try a few positive, skewed distributions for the histogram of clutch size:
```{r, message=FALSE, warning=FALSE}
hist(clutch$Clutch_size, 20, freq=F)
xx <- 0:15
fit_gamma <- fitdistr(clutch$Clutch_size, dgamma, list(shape=1))
points(dgamma(xx, fit_gamma$estimate[1]), pch=1)

fit_lognorm <- fitdistr(clutch$Clutch_size, dlnorm, list(meanlog=0, sdlog=1))
points(dlnorm(xx, fit_lognorm$estimate[1], fit_lognorm$estimate[2]), pch=2)

fit_rayleigh <- fitdistr(clutch$Clutch_size, drayleigh, list(scale=1))
points(drayleigh(xx, fit_rayleigh$estimate[1]), pch=3)

```

Flat bootstrapping at the species level
```{r, message=FALSE, warning=FALSE}
N <- dim(clutch)[1]
N_reps <- 500
bootstrap_means <- rep(0, N_reps)

for (i in 1:N_reps) {
  bootstrap_means[i] <- mean(sample(clutch$Clutch_size, N, replace=T))
}
hist(bootstrap_means, 20)
mean(bootstrap_means)
```

Bootstrapping from families.  This produces the right mean value, but a very non-normal looking histogram.

```{r, message=FALSE, warning=FALSE}
N_reps <- 500
family_means <- rep(0, N_reps)
for (i in 1:N_reps) {
  family <- sample(clutch$Family, 1)
  family_means[i] <- mean(clutch$Clutch_size[clutch$Family == family])
}
hist(family_means, 20)
mean(family_means)
```

We had some disagreement/confusion in our group as to what exactly we were supposed to do at the family level, so I did this another way. 

```{r, message=FALSE, warning=FALSE}
N_families <- length(unique(clutch$Family))
family_means_2 <- rep(0, N_reps)
these_means <- rep(0, N_families)

for (i in 1:N_reps) {
  for (j in 1:N_families) {
    family <- sample(clutch$Family, 1)
    these_means[j] <- mean(clutch$Clutch_size[clutch$Family == family])
  }
  family_means_2[i] <- mean(these_means)
}
hist(family_means_2, 20)
mean(family_means_2)

```
