Title Rebecca and Chloe's Week 2 Assignment
========================================================

```{r} 
species<-read.csv("/Users/rebeccakulp/Documents/Classes/2014/Stats II/PermuteSeminar-2014/Week-2/ClutchSize.csv",head=T)
hist(species$Clutch_size)
```
Distribution looks like a poisson or gamma distribution, so will need to test the two different distributions.

```{r} 

library(MASS)
species.poisson<-round(species$Clutch_size)
Poisson<-fitdistr(species.poisson,"Poisson")
gamma<-fitdistr(species.poisson,"gamma")
pois<-hist(rpois(5000,3.44))
gam<-hist(rgamma(5000,shape=3.99,rate=1.16))
```
Will graph the three different graphs on top of each other, so can see how the different distributions compare to the originial histogram.
```{r}
hist(species.poisson,col="yellow")
plot(gam,col="red", add=T)
plot(pois,col="blue", add=T)
```
The Poisson distribution matches the rounded distribution of the raw clutch distribution.

Now on to determinig how to bootstrap with entire dataset, and not considering the heirarchy.

```{r}
bootstrap.species<-matrix(nrow=1000,ncol=1)
for (i in 1:1000){
  d<-sample(species.poisson,2392,replace=T)
  e<-mean(d)
  bootstrap.species[i]<-e
}
overall.mean.species<-mean(bootstrap.species)
```

Now bootstrap, if first selecting from families.
```{r}
bootstrap.species<-sample(species.poisson,2392,replace=T)
families<-unique(species$Family)
length(families)
length(species.poisson)
```
Working code
```{r}
column<-cbind(species$Family,species$Clutch_size)
family.matrix<-matrix(1000,2)

```