Week 4 Notes
========================================================

fixing the problem of reading files(such as .csv) to Rmarkdown on Github
Install the RCurl package

```{r}
library (RCurl)
```
getUrL()  #use raw file

Read in data downloaded from blackboard  
```{r}
require(RCurl)
raw <- getURL("https://raw.github.com/PermuteSeminar/PermuteSeminar-2014/master/Week-2/ClutchSize.csv")
clutch <- read.csv(text = raw)

```


Lecture Notes
==============
Bootstrapping Time Series
-------------------------

use data without knowing the underlying proability distircution
use emperical distribution to understand distribution of statistic(mean, variance, etc)
sample with replacement from original * indicate resampled item.  
resampled population should equal number of samples

use standard deviation of replications to estimate standard error

emperical distribution  to recreate distribution

figure 8.1 and 8.3
Probability models may  be a whole process that involve multiple distributions 

Today's example:
------------------
Leutenizing hormone
time series with some temporal stucture
the whole distirbution may be normal
AR1 process  (type in equation here with Latex**)

Excercies
Use dataset from book :

1. Find beta (hat) best value that minimizes error
2. Sample with replacement from random deviations or do a moving blocks bootstrap (with replacement chunks of the data)




Get in touch with Heather about working with your own datasets*

Read Ch 15 and handouts, and paper for next week!

