Week 5 Emily Rollinson
========================================================
Readings for week 5:
Efron and Tibshirani, Chapter 15 "Permutation Tests"
Handout - Chapter 1 "Randomization" from Manly (2006)
Beijder et al. (1998)
Manly (1995)

In-class discussion
--------------------------------------------------------

In-class exercises
--------------------------------------------------------

```{r} 
#code for directly importing data from github
require(RCurl)
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
raw <- getURL("insertURLhere") #insert the  raw URL for the data file on github here
data <- read.csv(text = raw) #read in the github file
```

Notes and practice from the readings
--------------------------------------------------------
**From the Manly (2006) handout - the randomization test of golden jackals data**

Mandible lengths (mm) for male and female golden jackals.  Sample size of 10 for each sex.
```{r}
m <- c(120, 107, 110, 116, 114, 111, 113, 117, 114, 112) #male mandible lengths
f <- c(110, 111, 107, 108, 110, 105, 107, 106, 111, 111) #female mandible lengths
c <- c(m,f) #combined vector of male and female mandible lengths
obd <- mean(m) - mean(f) #observed difference between males and females

n = 10
perm = NULL
mtest = NULL
for (i in 1:9999){ #randomly generating 1000 new male and female data sets from the pooled original data and generating 1000 test statistics - the difference in means between the two groups)
  mper = sample(c, n, replace=FALSE)
  fper = sample(c, n, replace = FALSE)  #this is wrong, I need to include some dependence on the line above
  diff = (mean(mper)-mean(fper))
  perm = append(perm, diff)
}

hist(perm)
sig <- (sum((perm > obd)/9999))
sig
```

**Extraneous handy bits of code**
```{r}
#combining two vectors
v1 <- c(1,2,3)
v2 <- c(4,5,6)
v3 <- c(v1,v2)
#make it a data frame
V3 <- as.data.frame(v3)
#add a column
z <- c("a", "a", "a", "b", "b", "b")
V3$z = z
```{r}