Week 5 Emily Rollinson
========================================================
Readings for week 5:
Efron and Tibshirani, Chapter 15 "Permutation Tests"
Handout - Chapter 1 "Randomization" from Manly (2006)
Beijder et al. (1998)
Manly (1995)

In-class discussion
--------------------------------------------------------
To read later
"serial method" in Manly (1995)

In-class exercises
--------------------------------------------------------
Data are from Bejder et al. (1998)

Calculate HWI for the dolphin data
Write code for pairwise swapping
Examine the difference between generating null distributions by long chains vs. short chains; look for burn-in effect

```{r} 
#code for directly importing data from github
require(RCurl)
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
raw <- getURL("https://raw.github.com/PermuteSeminar/PermuteSeminar-2014/master/Week-5/Dolphin+data.csv") #insert the  raw URL for the data file on github here
dolphindata <- read.csv(text = raw) #read in the github file
```


HWI = x/ (x+y{ab}+ 0.5(y{a}+y{b}))
x = number of encounters including both dolphins A and B in the same group
ya = number of encounters including A but not B
yb = vice versa
yab = number of encounters including dolphins A and B, but in different groups (always 0 in this data set because different groups were not scored simultaneously)

Calculate HWI for each dyad from the observed data, then calculate a summary statistic for testing the randomness of co-occurrences

Calculating HWI
```{r}

x=matrix(nrow=18,ncol=18)
for(i in 1:18){
  for(j in 1:18)
    x
}



```



Permuting the original data
```{r}
mat1=matrix(nrow=2, data=c(0,1,1,0))
mat2=matrix(nrow=2, data=c(1,0,0,1))
```

```{r}

S <- sum(sum())
```