Week 13 Summary

========================================================
Based on data from Svoboda et al. 2014
========================================================

This week we explored questions similar to those posed in week 11.
We compared the communities of Desmids (a type of algae) in wetland communities.

Our primary goal was to test for spatial autocorrelation between the community composition of sites.

This autocorelation can be tested using the test statistic R given in Clark 1993.

$$
R= \frac{(\bar{r}_{b} - \bar{r}_{w})}{M/2}
$$

R is calculated based on the distances between points in a Multi Dimentional Scaling (MDS) analysis. There are several problems with using the MDS distances in a test statistic however.  Different numbers of dimentions in the MDS would result in different values for the test statistic, even though the number of dimensions is somewhat arbitrary.
These problems are solved by using the rank similarities between samples.  More similar samples are ranked higher than less similar samples.  Thus, in the test statistic R, $\bar{r}_{b}$ gives the average rank of the differences between sites while $\bar{r}_{w}$ gives the average rank of differences within sites. M is simply a constant based on sample size that scales R so that it does not exceed the range -1:1.

Excercise
=================
The data provided were transformed in several different ways, and different analyses may require the data in different forms from what I will be using here, but the files I will be using are pre-transformed and uploaded on GitHub.

```{r}
require(RCurl)
require(vegan)
require(ggplot2)
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
      
raw <- getURL("https://raw.githubusercontent.com/PermuteSeminar/PermuteSeminar-2014/master/Week-13/Svoboda_transpose.csv")
data <- read.csv(text = raw, row.names=1) 
raw.2 <- getURL("https://raw.githubusercontent.com/PermuteSeminar/PermuteSeminar-2014/master/Week-13/EJR_siteinfo.csv")
Groups <- read.csv(text=raw.2)
```

First we can visualize the data in several different ways

```{r}
cca.data<-cca(data)
plot(cca.data) 
# The crosses are species and the circles are sites


MDS.data<-metaMDS(data, distance="bray", trymax=50)
points<-as.data.frame(MDS.data$points)
points <- cbind(points, type = Groups$Type)
ggplot(points,aes(x=MDS1, y=MDS2, col=type)) + geom_point()
#this is color-coded for bogs vs. fens, you could color code the pools as well
```


Now that we have visualized the data it is pretty clear that there is a difference between the Bog and Fen sites.
To statistically test the differences between the two habitats or the sites within these habitats we need to calculate R, and then permute the similarities in order to create a bootstrap distribution to test the significance of our test statistic. 
We could do this all by hand, but luckily the vegan package in R can do it all for us. 
The function anosim both uses R as its test statistic, and it provides a p value via permutation.

So to compare our sites all we have to do is the following.
```{r}
#Difference between fen and bog
anosim(data, Groups$Type, permutations=999, distance="bray")

#Differences among fens
anosim(data[1:9,], Groups[1:9,]$Site, permutations=999, distance="bray")

#Differences among bogs
anosim(data[10:18,], Groups[10:18,]$Site, permutations=999, distance="bray")
```

All three comparisons show significant differences meaning that, not only are the communities of desmids in bogs and fens different, each site has a different community composition.  This means that community composition of desmids is strongly spatially autocorrelated and not well explained by differences in abiotic factors of the water.