Permute Seminar, Week 1
========================================================

Load the libraries that we will need for examples
```{r}
library(ggplot2)
```

**For Week 2:** read chapters 3-5

**Notes:**
Bootstrap = Sample with replacement

Boostrapped datasets are generally represented using a "*".
Estimates are represented using "^".

Sampling with replacement from original data can take the place of redoing an experiment many times.

The standard deviation of the distribution of the bootstrapped estimate is the standard error of the statistic.
Can pull the confidence intervals from the percentiles of this distribution.

iid = independent and identically distributed

**From the text:**
Chapter 1
----------

Aspirin study group data
```{r}
aspirin = c(rep(1,119), rep(0,10918))
placebo = c(rep(1,98), rep(0,10936))
```

obtain 1000 bootstrap replicates for theta using random sampling with replacement
```{r}
est=NULL
for (i in 1:1000){
  bootA = sample(aspirin, 11037, replace=TRUE)
  bootB = sample(placebo, 11034, replace=TRUE)
  theta = (sum(bootA)/length(bootA))/(sum(bootB)/length(bootB))
  est = append(est, theta)
}

hist(est, xlab = "boostrap estimated means")
```

# plot and summarize the results
```{r}
se <- function(x) sqrt(var(x)/length(x))

hist(est)
mean(est)
sd(est)
se(est)
```

# find the confidence intervals
```{r}
quantile(est, 0.95)
quantile(est, 0.05)
```

# Chapter 2

#mouse study data, lifespan (days) after surgical treatment
```{r}
trmt = c(94, 197, 16, 38, 99, 141, 23)
control = c(52, 104, 146, 10, 51, 30, 40, 27, 46)

mean(trmt)
mean(control)
diff = mean(trmt) - mean(control)

se(trmt)
se(control)
sediff = sqrt(se(trmt)^2 + se(control)^2)
```

Thus, the observed difference is only about 1 stderr greater than 0; Insiginificant.
```{r}
diff/sediff
```

Compare the sample medians
```{r}
median(trmt)
median(control)
diff2 = median(trmt) - median(control)
```

Obtain 1000 bootstrap replicates for mean using random sampling with replacement
```{r}
meanA=NULL
meanB=NULL
medianA=NULL
medianB=NULL
for (i in 1:200){
  bootA = sample(trmt, 7, replace=TRUE)
  bootB = sample(control, 9, replace=TRUE)
  meanA = append(meanA, mean(bootA))
  meanB = append(meanB, mean(bootB))
  medianA = append(medianA, median(bootA))
  medianB = append(medianB, median(bootB))
}

seBootA = sd(meanA)
seBootB = sd(meanB)
seBootAm = sd(medianA)
seBootBm = sd(medianB)
```

The observed difference in means is less than 1 sd greater than 0
```{r}
sediff = sqrt(seBootA**2 + seBootB**2)
diff/sediff
```

The observed difference in medians is greater than the difference in means, but is still insignificant.
```{r}
sediff2 = sqrt(seBootAm**2 + seBootBm**2)
diff2/sediff2
```

Problems:
#2.1 Supoose mouse survival was in weeks instead of days. How would the answers change?
```{r}
trmt = trmt/7
control = control/7
```

