Emily Rollinson Week 13
========================================================

Data from Svoboda et al. 2014 - Spatial and temporal community structure of desmids

Load packages
```{r}
require(RCurl)
require(vegan)
```

Load data
```{r}
comm<-read.csv("Svoboda_transpose.csv")
groups<-read.csv("EJR_siteinfo.csv")
```

Run NMDS - plots the three different sites

```{r}
ord<-metaMDS(comm[,2:96], distance="bray", k=3, trymax=50)
plot(ord, display="sites")
```

Run anosim on the various groupings

```{r}
anosim(comm, groups$Type, permutations=999, distance="bray")
anosim(comm, groups$Site, permutations=999, distance="bray")
```


----------------------
Team Swamp Code below
----------------------

Methods from Clark 1993 - section "A permutation test for the one-way layout: Frierfjord macrobenthos".
rw = average of all rank similarities among replicates within sites
rb = average of rank similarities among all pairs of replicates between different sites
M = n((n-1))/2; n = total $ of samples being considered

Our goal is to examine site to site differences in community structure (in this case, differences among the pools within the swamp sites - other groups are looking at differences among sites and differences among pools within the bog sites). 

Calculating rank similarities
```{r}
bray<-vegdist()

```


```{r}
# Transpose table YOU WANT
fooData.T <- t(fooData[,2:ncol(fooData)])
sp_t <- t(species[,2:ncol(species)])

# Set the column headings
colnames(fooData.T) <- fooData[1,]
colnames(sp_t) <- species[1:nrow(species),]

```
```