Permute Seminar, Week 1
========================================================

**For Week 2:** read chapters 3-5

**Notes:**
Bootstrap = Sample with replacement

Boostrapped datasets are generally represented using a "*".
Estimates are represented using "^".

Sampling with replacement from original data can take the place of redoing an experiment many times.

The standard deviation of the distribution of the bootstrapped estimate is the standard error of the statistic.
Can pull the confidence intervals from the percentiles of this distribution.

iid = independent and identically distributed

**Examples:**
Load the libraries that we will need for examples
```{r}
library(ggplot2)
```

Lets write a few functions that can be useful for us later today:

Make a dataset with a mean of 1 and sd of 3
Obtain 1000 bootstrap replicates using random sampling with replacement and take the mean of each
```{r}
data = rnorm(1000, 1, 3)
mean(data)

est=NULL
for (i in 1:1000){
  boot = sample(data, 1000, replace=TRUE)
  mean = mean(boot)
  est = append(est, mean)
}
mean(est)
sd(est)
```

Obtain 1000 datasets, and compare the means
```{r}
est2=NULL
for (i in 1:1000){
  data = rnorm(1000, 1, 3)
  mean = mean(data)
  est2 = append(est2, mean)
}
mean(est2)
sd(est2)
hist(est2)
```

Compare the histograms directly. They won't always look the same, because they move around somewhat.
```{r}
par(mfrow=c(1,1))
hist(est, col = "cadetblue", xlim=c(0.5,2))
hist(est2, col="indianred", add=TRUE)
```

2. Draw 1000 from a uniform distribution with mean of zero and sd of 5
Find the maximum value. 
What is the probability that the maximum value in your dataset is also the largest number in the bootstrapped dataset?
When sample from the original distribution, you are sampling from the continuous distribution
When sample with replacement 
```{r}
data = runif(1000, 0, 5)
max(data)

est=NULL
for (i in 1:1000){
  boot = sample(data, 1000, replace=TRUE)
  maxi = max(boot)
  est = append(est, maxi)
}

max(est)
sd(est)

est2=NULL
for (i in 1:1000){
  data = runif(1000, 0, 5)
  maxi = max(data)
  est2 = append(est2, maxi)
}
max(est2)
sd(est2)

hist(est, col = "cadetblue")
hist(est2, col="indianred", add=TRUE)
```
1000 samples, 1 maximum value
Probability of drawing max value is 1/n
Probability of not drawing max value on first draw: 1 - (1/n)
You have to AVOID the biggest value n times: 1 - (1 - (1/n))^n
As n goes to infinity, this value is e^-1
```{r}
1-(1-(1/1000))^1000
```
Thus, 63% of the time, the largest value is the same as what you had in the original dataset in your bootstrap.Using the other method, you get an exponential distribution. 

When looking at extreme values (min, max) or median, where a relatively small subset of values are going to keep showing up, it can be a bad decision to use the bootstrap. Absolute values can cause discontinuity, or if the underlying values are not continuous (e.g., my movement data?)

The **Cauchy distribution** is "pathological" and has sub-exponential tails.
It does not play well with the bootstrap.

Lets try the same thing as above, but lets use the variance as the test statistic:
```{r}
data = rcauchy(1000, 1, 2)
mean(data)

est=NULL
for (i in 1:1000){
  boot = sample(data, 1000, replace=TRUE)
  vari = var(boot)
  est = append(est, vari)
}

est2=NULL
for (i in 1:1000){
  data = rcauchy(1000, 1, 2)
  vari = var(data)
  est2 = append(est2, vari)
}
mean(est2)
mean(est)
sd(est2)
sd(est)

hist(est2, col = "cadetblue") #spectacular failure
hist(est, col="indianred")

```


**From the text:**
Chapter 1
----------

Aspirin study group data
```{r}
aspirin = c(rep(1,119), rep(0,10918))
placebo = c(rep(1,98), rep(0,10936))
```

obtain 1000 bootstrap replicates for theta using random sampling with replacement
```{r}
est=NULL
for (i in 1:1000){
  bootA = sample(aspirin, 11037, replace=TRUE)
  bootB = sample(placebo, 11034, replace=TRUE)
  theta = (sum(bootA)/length(bootA))/(sum(bootB)/length(bootB))
  est = append(est, theta)
}

hist(est, xlab = "boostrap estimated means")
```

# plot and summarize the results
```{r}
se <- function(x) sqrt(var(x)/length(x))

hist(est)
mean(est)
sd(est)
se(est)
```

# find the confidence intervals
```{r}
quantile(est, 0.95)
quantile(est, 0.05)
```

# Chapter 2

#mouse study data, lifespan (days) after surgical treatment
```{r}
trmt = c(94, 197, 16, 38, 99, 141, 23)
control = c(52, 104, 146, 10, 51, 30, 40, 27, 46)

mean(trmt)
mean(control)
diff = mean(trmt) - mean(control)

se(trmt)
se(control)
sediff = sqrt(se(trmt)^2 + se(control)^2)
```

Thus, the observed difference is only about 1 stderr greater than 0; Insiginificant.
```{r}
diff/sediff
```

Compare the sample medians
```{r}
median(trmt)
median(control)
diff2 = median(trmt) - median(control)
```

Obtain 1000 bootstrap replicates for mean using random sampling with replacement
```{r}
meanA=NULL
meanB=NULL
medianA=NULL
medianB=NULL
for (i in 1:200){
  bootA = sample(trmt, 7, replace=TRUE)
  bootB = sample(control, 9, replace=TRUE)
  meanA = append(meanA, mean(bootA))
  meanB = append(meanB, mean(bootB))
  medianA = append(medianA, median(bootA))
  medianB = append(medianB, median(bootB))
}

seBootA = sd(meanA)
seBootB = sd(meanB)
seBootAm = sd(medianA)
seBootBm = sd(medianB)
```

The observed difference in means is less than 1 sd greater than 0
```{r}
sediff = sqrt(seBootA**2 + seBootB**2)
diff/sediff
```

The observed difference in medians is greater than the difference in means, but is still insignificant.
```{r}
sediff2 = sqrt(seBootAm**2 + seBootBm**2)
diff2/sediff2
```

Problems:
#2.1 Supoose mouse survival was in weeks instead of days. How would the answers change?
```{r}
trmt = trmt/7
control = control/7
```


